---
title: "Introduction to varstan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro-to-varstan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

Varstan is a a package for bayesian estimation of structured time series models,using a Hamiltonian monte carlo, implemented with the so popular package "rstan". The aim of varstan is to have an interface of the most popular time series modeles such as: arima, garch, sarima, stochastic Volatility models (SVM), Hiden Markov models(HMM), seasonal fouier regresión, additive non-linear models (*via prophet package*), univariate kalman Filters, varma and bekk models.

On the beta version 0.5.0.000, the avaliable models are:
 
  + arima
  + garch
  + varma
  + Generalized t-student varma

The dynamic of varstan is to build your own model using one of the avaliable model constructor, personalize your own priors (*check the Use Priors vignette*), and fit your model using the varstan function. On the next example we show you how to create and fit a simple bayesian arima model.


```{r setup}
library(bayesplot)
library(varstan)
library(rstan)
library(forecast)
library(ggplot2)
library(gridExtra)
```

First step is make a simulation of a simple arma model with 200 observations as follows:

$$Y_t = \mu_0 + 0.338Y_{t-1} - 0.2279\epsilon_{t-1} - 0.2488\epsilon_{t-2}, \text{ } \epsilon_t \sim N(0,\sigma^2_0)  $$

```{r}
y = arima.sim(n = 700, list(ar =0.35, ma = c(0.2279, 0.2488)),sd = sqrt(0.1796))

autoplot(y)+labs(x = "time",title = "Simulated ARMA Process")+theme_classic()
```

Proceding to built the arima model using the varstan constructor:
```{r}
model1 = arima(y,p = 1,d = 0,q = 2)
```

Automaticly varstan build a bayesian arima model, with default normal priors, you can check the model using the report function:

```{r}
report(model1)
```

Changing the prior of one of the defined parameter, just use the set_prior function, in this example we are gonna change the second ma component for a beta distribution

$$\theta_2 \sim beta(2.5,2.5)$$

```{r}
model1 = set_prior(model1,type = "ma",par1 = 2.5,par2 = 2.5,lag = 2,dist = "beta")
get_prior(model1,type = "ma")
```

To see more details of the avaliable priors and the model structure see the vignettes *Use_prior* and *arima_models*, respectively. For estimating the defined model, use the varstan function, it will estimate the posterior sample using a Hamiltonian montecarlo.

In this example a hmc is run with 1 chain of 4000 iterations 

```{r}
sfit = varstan(model1,chains = 1,iter = 7000)
```


```{r}
summary(sfit,robust= TRUE)
```


### Parameter Diagnostic

We can see the whole posterior distribution and make some diagnostics

```{r}
post = as.data.frame(extract(sfit$stanfit,"phi", permuted = TRUE) )
  color_scheme_set("viridis")
  p1 = mcmc_trace(post,  pars = "phi",
        facet_args = list(nrow = 2, labeller = label_parsed)) + 
        facet_text(size = 15)+theme_classic()
  p2 = mcmc_hist(post, pars = "phi",facet_args = list(nrow = 2))+
    facet_text(size = 15)+theme_classic()
  p3 = mcmc_acf(post, pars = "phi", lags = 10,)+theme_minimal()
  grid.arrange(p1,p2,p3,nrow = 2,layout_matrix = matrix(c(1,3,2,3),ncol=2,byrow=TRUE))

```



# The classical arima estimation is:

```{r}
mc = stats::arima(y,order = c(1,0,2))
mc
```


The residual values of the varstan object is

```{r}
resid = residuals(sfit)
```


The residuals of the classical estimation are:

```{r}
summary(mc$residuals)
```

The posterior median residual statistics of the bayesian model are:
```{r}
summary(resid)
```

