---
title: "Introduction to varstan"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Intro-to-varstan}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette is used to show varstan's dynamic for estimating a time series model, in this case we are gonna use the data *Astrovan*, a simulated var(1)bekk(1,1) model, available in the varstan package.

The fist part is to call the varstan library 
```{r setup}
library(varstan)
```

For this example, a var(1)-bekk(1,1) model is used for the data, the model is defined with the model constructure: **varbekk**,Where p,s and k  are the lags of the var, arch and garch components.
```{r}
dat1 = varbekk(ts = Astrovan,p = 1,s = 1,k = 1)
```

Varstan already use default normal priors to check them just print the model object:
```{r}
print(dat1)
```

The set_prior function is available for changing the prior hyperparameters as follow:

```{r}
dat1 = set_prior(dat1,parameter = "var",location = 0,scale = 5)
print(dat1)
```

The varstan function is used to estimating the current model, it returns a varstan object that has 2 important parts:

 + the returned stanfit object
 + the user defined model.
 
```{r,eval = FALSE}
fit1 = varstan(dat1,t.student = FALSE,chains = 1,iter = 1000,adapt.delta = 0.9)
```

The summary of the estimated parameters is:

```{r,eval = FALSE}
print(fit1)
```

